### 2015-11-10 paper/pldi16-submission/evaluation.tex
texing

\begin{figure}
\judgementHead{Evaluation}{\stepsto{e, \su}{e, \su}}
$$
\inference[\recontext]{\stepsto{e, \su}{e_1, \su_1}}{\stepsto{C[e], \su}{C[e_1], \su_1}}
$$

$$
\inference[\restuck]{}{\stepsto{C[\stuck], \su}{\stuck, \su}}
$$

$$
\inference[\replusgood]
  {n_1, \su_2 = \force{v_1}{\tnat} \\
   n_2, \su_3 = \force{v_2}{\tnat} \\ 
   n = \eplus{n_1}{n_2}}
  {\stepsto{\eplus{v_1}{v_2}, \su_1}{n, \su_1\su_2\su_3}}
$$

$$
\inference[\replusbadone]
  {\stuck, \su_2 = \force{v_1}{\tnat}}
  {\stepsto{\eplus{v_1}{v_2}, \su_1}{\stuck, \su_1\su_2}}
$$

$$
\inference[\replusbadtwo]
  {\stuck, \su_2 = \force{v_2}{\tnat}}
  {\stepsto{\eplus{v_1}{v_2}, \su_1}{\stuck, \su_1\su_2}}
$$

$$
\inference[\reifgoodone]
  {\etrue, \su_2 = \force{v}{\tbool}}
  {\stepsto{\eif{v}{e_1}{e_2}, \su_1}{e_1, \su_1\su_2}}
$$

$$
\inference[\reifgoodtwo]
  {\efalse, \su_2 = \force{v}{\tbool}}
  {\stepsto{\eif{v}{e_1}{e_2}, \su_1}{e_2, \su_1\su_2}}
$$

$$
\inference[\reifbad]
  {\stuck, \su_2 = \force{v}{\tbool}}
  {\stepsto{\eif{v}{e_1}{e_2}, \su_1}{\stuck, \su_1\su_2}}
$$
\caption{Evaluation relation}
\label{fig:evaluation}
\end{figure}