### 2015-11-11 paper/pldi16-submission/main.tex
texified proof sketch

%
% LaTeX template for prepartion of submissions to PLDI'15
%
% Requires temporary version of sigplanconf style file provided on
% PLDI'15 web site.
% 
\documentclass[pldi,indentedstyle=false]{sigplanconf-pldi15}

%
% the following standard packages may be helpful, but are not required
%
\usepackage{amsmath}
\usepackage{amssymb}
% \usepackage{SIunits}            % typset units correctly
\usepackage{courier}            % standard fixed width font
\usepackage[scaled]{helvet} % see www.ctan.org/get/macros/latex/required/psnfss/psnfss2e.pdf
\usepackage{url}                  % format URLs
\usepackage{listings}          % format code
\usepackage{enumitem}      % adjust spacing in enums
\usepackage[colorlinks=true,allcolors=blue,breaklinks,draft=false]{hyperref}   % hyperlinks, including DOIs and URLs in bibliography
% known bug: http://tex.stackexchange.com/questions/1522/pdfendlink-ended-up-in-different-nesting-level-than-pdfstartlink
\newcommand{\doi}[1]{doi:~\href{http://dx.doi.org/#1}{\Hurl{#1}}}   % print a hyperlinked DOI
\usepackage{xspace}
\usepackage{commands}

\usepackage{amsthm}
% \newtheorem{lemma}{Lemma}
% \newtheorem{corollary}{Corollary}
% \newtheorem{theorem}{Theorem}
% \newtheorem{definition}{Definition}
\theoremstyle{plain}% default 
\newtheorem{thm}{Theorem} % [section] 
\newtheorem{lem}[thm]{Lemma} 
\newtheorem{prop}[thm]{Proposition} 
\newtheorem*{cor}{Corollary} 
\theoremstyle{definition}
\newtheorem{defn}{Definition} % [section]

\usepackage[inference]{semantic}

\begin{document}

%
% any author declaration will be ignored  when using 'plid' option (for double blind review)
%

\title{NanoMaLy}

\authorinfo{Eric L. Seidel\and Ranjit Jhala}
           {UC San Diego}
           {{eseidel,rjhala}@cs.ucsd.edu}

\authorinfo{Westley Weimer}
           {University of Virginia}
           {weimer@viginia.edu}

\maketitle
\begin{abstract}
\end{abstract}

% 11 pages total (not including bib)
% ----------------------------------
%  1p : intro
%  2p : overview
%  2p : type-carrying semantics
%  1p : evaluation: quickchecking type errors
%  2p : interactive semantics
%  1p : evaluation: nanomaly
%  1p : related work
%  1p : conclusion

\section{Introduction}          % 1 page
\section{Overview}              % 2 pages?
\section{Type-carrying Semantics} % 2 pages
\input{semantics.tex}

\section{Evalution: Recasting Type Errors as Runtime Errors} % 1 page
\label{sec:eval-recast-type}
\begin{itemize}
\item benchmarks: our data + seminal data
\item both cases: \textbf{random} search sufficient to trigger runtime crash in 80\% of programs
\item how many of the ``safe'' programs are actually safe??
\end{itemize}

\section{``Interactive'' Semantics \& Trace Exploration} % 2 pages
\begin{itemize}
\item extend operational semantics to collect reduction graph
\item nodes are terms, edges indicate ``steps-to'' and ``sub-term'' relations
\item visualize path through reduction graph
\item expand edges to reveal more fine-grained steps (step/jump forward/backward)
\item never lose context (unlike traditional debugger)
\end{itemize}

\section{Evaluation: NanoMaLy}                % 1 page
\section{Related Work}              % 1 pages

\subsection{Type Error Localization}
\label{sec:type-error-local}
\begin{itemize}
\item seminal
\item bayesian (Zhang+Myers)
\item MaxSMT
\item counterfactual (Chen+Erwig)
\end{itemize}
\subsection{Debugging \& Program Understanding}
\label{sec:debugging}

\section{Conclusion}                % 1 page

\end{document}