### 2015-11-18 paper/pldi16-submission/related.tex

\section{Related Work}
\label{sec:related-work}

\subsection{Type Error Localization}
\label{sec:type-error-local}
\begin{itemize}
\item \cite{lerner_searching_2007}
\item \cite{zhang_toward_2014}
\item \cite{pavlinovic_finding_2014}
\item \cite{chen_counter-factual_2014}
\end{itemize}

\subsection{Testing}
\label{sec:testing}
\begin{itemize}
\item \cite{claessen_quickcheck:_2000}
\item \cite{runciman_smallcheck_2008}
\item \cite{godefroid_dart:_2005}
\item \cite{cadar_klee:_2008}
\item \cite{tillmann_pexwhite_2008}
\end{itemize}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: