### 2015-11-19 paper/pldi16-submission/evaluation.tex
consolidate evaluation sections into one

% \section{Evaluation: Recasting Type Errors as Runtime Errors}
\section{Evaluation}
\label{sec:evaluation}

%
The immediate question is ``What fraction of type errors admit
witnesses?''
%
To answer this question we implemented a prototype of our search
procedure for the pure subset of \ocaml, \ie \lang extended with
algebraic datatypes and records. 
%
In our implementation we instantiated \gensym with a simple random
generation of values, which we will show is more than sufficient for the
majority of type errors.
%
We evaluated our implementation of two sets of known-bad programs, \ie
programs that were rejected by the \ocaml compiler because of a type
error.
%
The first dataset comes from the Spring 2014 undergraduate Programming
Languages course at UC San Diego (IRB \#140608). 
%
We recorded each interaction with the \ocaml top-level system over the
course of the first three assignments, from which we extracted 4,407
distinct, ill-typed \ocaml programs.
%
The second dataset -- widely used in the literature -- comes from a
similar course at the University of
Washington~\cite{lerner_seminal:_2006}, from which we extracted 284
ill-typed programs.

\subsection{Witness Coverage}
\label{sec:eval:witness-coverage}
%
We ran our search algorithm on each program for 1,000 iterations, with
the entry point set to the function that \ocaml had identified as
containing a type error.
%
Due to the possibility of non-termination we set a timeout of 10 minutes
per program.
% Due to the possibility of non-termination we set a limit on the number
% of reductions to perform, increasing in 500-step increments from 500
% steps to 3,000 steps total.
%
We also added a na\"ive check for infinite recursion; at each recursive
function call we check whether the new arguments are identical to the
current arguments.
%
If so, the function cannot possibly terminate and we report an error.
%
While not a \emph{type error}, infinite recursion is still a clear bug
in the program, and thus valuable feedback for the user.

\begin{figure}[t]
\centering
\begin{minipage}{\linewidth}
\includegraphics[width=\linewidth]{coverage.pdf}
\end{minipage}
\begin{minipage}{\linewidth}
\includegraphics[width=\linewidth]{distrib.pdf}
\end{minipage}
% \vspace{-8ex}
\caption{Results of our coverage testing and the distribution of test
  outcomes. Our random search successfully finds witnesses for over 85\%
  of the programs at the lowest timeout and improves slightly to 88\% at
  the next level. In both datasets we detect actual type errors about
  82\% of the time, unbound variables or constructors 3-4\% of the time,
  and diverging loops 2-3\% of the time. For the remaining 11-12\% of
  the programs we are unable to provide any useful feedback.  }
\label{fig:results-witness}
\end{figure}

\paragraph{Results}
\label{sec:results-witness}
The results of our experiments are summarized in
Figure~\ref{fig:results-witness}.
%
In both datasets our tool was able to find a witness for at least 87\%
of the programs in under 10 seconds.
%
Interestingly, while the vast majority of witnesses corresponded to a
type-error, as expected, 3-4\% triggered an unbound variable error and
2-3\% triggered an infinite recursion error.
%
For the remaining 12\% of programs we were unable to provide any useful
feedback as they either completed 1,000 tests successfully, or timed out
after 10 minutes.
%
% XX programs were deemed safe and XX timed out even at 3,000 steps, \ie
% we could not provide any useful feedback for XX\% of the total programs.
%
While a more advanced search procedure, \eg dynamic-symbolic execution,
could likely trigger more of the type errors, our experiments show that
type errors are coarse enough (or that novice programs are \emph{simple}
enough) that these techniques are not necessary.

\subsection{Trace Complexity}
\label{sec:trace-complexity}
For each of the ill-typed programs for which we could find a witness, we
measure the complexity of the generated trace according to the following
metrics:
%
\begin{enumerate}
\item The number of single-step reductions along the path from the
  witness to the stuck term. This can be thought of as a worst-case
  complexity, \ie ``How big is the fully-expanded trace?''
\item The number forward (or backward) jumps that can be taken along the
  path from the witness to the stuck term. This is not quite the same as
  a stack-trace as it shows \emph{all} function calls, including those
  that returned successfully.
  % These abstract many details of the computation that are often
  % uninteresting or irrelevant to the explanation.
% \item \ES{others?}
\end{enumerate}
%
\begin{figure}[t]
\centering
% \includegraphics[width=\linewidth]{trace_size.pdf}
% \begin{minipage}{0.49\linewidth}
\includegraphics[width=\linewidth]{trace_size_step.pdf}
% \end{minipage}
% \begin{minipage}{0.49\linewidth}
\includegraphics[width=\linewidth]{trace_size_jump.pdf}
% \end{minipage}
\caption{Complexity of the generated traces. At least 90\% of the traces
  have a jump complexity of less than 10, with an average of 3.}
\label{fig:results-complexity}
\end{figure}
%

\paragraph{Results}
\label{sec:results-complexity}
The results of the experiment are summarized in
Figure~\ref{fig:results-complexity}.
%
The average number of single-step reductions per trace is 26 for the
UCSD dataset (27 for the Seminal dataset) with a maximum of 2,744 (659
for Seminal). The average number of jumps per trace is 3 (also 3 for
Seminal) with a maximium of 177 (52 for Seminal). In both datasets at
least 90\% of the traces contain fewer than 10 jumps.
%

\subsection{Discussion}
\label{sec:discussion}
To summarize, our experiments demonstrate that \nanomaly finds witnesses
to type errors: (1) with high coverage in a timespan amenable to
compile-time analysis, and (2) with traces that have a low average
complexity of 3 function calls.

There are, of course, drawbacks to our approach.

\paragraph{Random Generation}
Random test generation has difficulty generating highly constrained
values, \eg red-black trees or a pair of equal integers. If the type
error is hidden behind a complex branch condition \nanomaly may not be
able to trigger it. Exhaustive testing and dynamic-symbolic execution
can address this short-coming by performing an exhaustive search for
inputs (\emph{resp}. paths through the program). As our experiments
show, however, novice programs do not appear to require more advanced
search techniques, likely because the novice programs tend to be simple.

\paragraph{Infinite Types}
Our algorithm does check for infinite types inside \forcesym, but there
are some degenerate cases where it is unable to detect them. Consider,
for example, the following buggy version of @replicate@ which is
supposed to produce a list containing @n@ copies of @x@.
%
\begin{code}
  let rec replicate n x =
    if n <= 0 then
      []
    else
      replicate (n-1) [x]
\end{code}
%
This code produces a nested list (with @n@ levels of nesting) containing
a single copy of @x@. \ocaml detects a cyclic \hbox{@'a = 'a list@}
constraint in the recursive call and throws a type error, whereas
\nanomaly happily recurses as many times as necessary. Strictly
speaking, this function itself cannot ``go wrong'', the program would
not get stuck until a \emph{client} attempted to use the result
expecting a flat list. But this is not very satisfying as @replicate@ is
clearly to blame. Furthermore, in our experience, infinite-type errors
are some of the more difficult ones to debug (and to explain to
novices), so better support for this scenario would be useful.

\paragraph{Trace Explosion}
Though the average complexity of our generated traces is low in terms of
jumps, there are some extreme outliers. We cannot reasonably expect a
novice user to explore a trace containing 50+ terms and draw a
conclusion about which pieces contributed to the bug in their
program. Enhancing our visualization to slice out program paths relevant
to specific values, \eg in the style of~ \cite{perera_functional_2012},
would likely help alleviate this issue. This would allow users to
highlight a confusing value and ask: ``Where did this come from?''

% \begin{itemize}
% \item benchmarks: our data + seminal data
% \item both cases: \textbf{random} search sufficient to trigger runtime crash in 80\% of programs
% \item how many of the ``safe'' programs are actually safe??
% \end{itemize}