### 2015-11-19 paper/pldi16-submission/main.tex
done with abstract

%
% LaTeX template for prepartion of submissions to PLDI'15
%
% Requires temporary version of sigplanconf style file provided on
% PLDI'15 web site.
%
\documentclass[blind,preprint,nocopyrightspace,10pt,squareay,clearpagebib,explicitsize]{sigplanconf-pldi15}

%
% the following standard packages may be helpful, but are not required
%
\usepackage{amsmath}
\usepackage{amssymb}
% \usepackage{SIunits}            % typset units correctly
\usepackage{courier}            % standard fixed width font
\usepackage[scaled]{helvet} % see www.ctan.org/get/macros/latex/required/psnfss/psnfss2e.pdf
\usepackage{url}                  % format URLs
\usepackage{listings}          % format code
\usepackage{enumitem}      % adjust spacing in enums
\usepackage[colorlinks=true,allcolors=blue,breaklinks,draft=false]{hyperref}   % hyperlinks, including DOIs and URLs in bibliography
% known bug: http://tex.stackexchange.com/questions/1522/pdfendlink-ended-up-in-different-nesting-level-than-pdfstartlink
\newcommand{\doi}[1]{doi:~\href{http://dx.doi.org/#1}{\Hurl{#1}}}   % print a hyperlinked DOI
\usepackage{xspace}
\usepackage{commands}
\usepackage{graphicx}
\usepackage{pdfpages}
%\usepackage[doi]{natbib}

\usepackage{amsthm}
% \newtheorem{lemma}{Lemma}
% \newtheorem{corollary}{Corollary}
% \newtheorem{theorem}{Theorem}
% \newtheorem{definition}{Definition}
\theoremstyle{plain}% default
\newtheorem{thm}{Theorem} % [section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem*{cor}{Corollary}
\theoremstyle{definition}
\newtheorem{defn}{Definition} % [section]

\usepackage{pgfplots}
\usepackage{pgfkeys}
\pgfplotsset{compat=1.12}

\usepackage[inference]{semantic}

\input{haskellListings}

\begin{document}

%
% any author declaration will be ignored  when using 'plid' option (for double blind review)
%

\title{NanoMaLy}

\authorinfo{Eric L. Seidel}
           {UC San Diego}
           {eseidel@cs.ucsd.edu}

\authorinfo{Ranjit Jhala}
           {UC San Diego}
           {jhala@cs.ucsd.edu}

\authorinfo{Westley Weimer}
           {University of Virginia}
           {weimer@viginia.edu}

\maketitle

\input{abstract}
\input{introduction}
\input{overview}
\input{semantics}
% \section{interactive}
\input{evaluation}
\input{related}

{
\bibliographystyle{abbrvnat}
\bibliography{main}
}

\end{document}