### 2015-11-20 paper/pldi16-submission/introduction.tex

\section{Introduction}
\label{sec:introduction}

Type errors are a common stumbling block for students
trying to learn typed functional languages like \ocaml\
and \haskell.
%
Consider the ill-typed @fac@ function on the left in
Figure~\ref{fig:factorial}.
%
The function returns @true@ in the base case (instead of @1@).
and so \ocaml's responds with the error message:
%
\begin{verbatim}
  File "fac.ml", line 5, characters 8-23:
  Error: This expression has type bool
         but an expression was expected
         of type int
\end{verbatim}

This message is makes perfect sense to an expert who is familiar
with the language and has a good mental model of how the type
system works.
%
However, it may perplex a novice who is still to develop such a
mental model.
%
To make matters worse, unification based type inference algorithms
often report errors far removed from their source.
%
This further increases the novice's confusion and can actively mislead
them to focus their investigation on an irrelevant piece of code.
%
Much recent work has focused on analyzing unification constraints
to properly \emph{localize} a type error~\cite{lerner_searching_2007,chen_counter-factual_2014,zhang_toward_2014,pavlinovic_finding_2014},
but an accurate source location still does not explain \emph{why} the
program is wrong.

\begin{figure}[t]
\centering
\begin{minipage}{.49\linewidth}
\begin{code}
let rec fac n =
  if n <= 0 then
    true
  else
    n * fac (n-1)
\end{code}
\end{minipage}
\hspace{0.3in}
\begin{minipage}{.3\linewidth}
  \includegraphics[height=2.5in]{fac-overview.png}
\end{minipage}
\caption{(L) An ill-typed \texttt{fac} function; (R) Dynamically witnessing the type error in \texttt{fac}}
\label{fig:factorial}
\end{figure}

In this paper we propose a new approach to explaining type
\emph{static} type errors: by \emph{dynamically} witnessing
how exactly an ill-typed program goes wrong.
%
We have developed \toolname\ an interactive tool which uses
the source of the ill-typed function to automatically synthesize
the result on the right in Figure~\ref{fig:factorial}, which
shows how the recursive calls reduce to a configuration where
the program ``goes wrong'' \ie\ the @int@ value @1@ is to be
multiplied with the @bool@ value @true@.
We achieve this via three concrete contributions.

\paragraph{1. Finding Witnesses}
Our first contribution is an algorithm for searching for
\emph{witnesses} to type errors, \ie\ inputs that cause a
program to go wrong~(\S~\ref{sec:semantics}).
%
This problem is surprisingly tricky when we cannot rely on
static type information.
%
In particular, we must avoid the trap of \emph{spurious} inputs
that cause irrelevant problems that would be avoided by picking
values of a different, relevant type.
%
We solve this problem by developing a novel operational semantics
that corresponds to executing the program with \emph{holes}
values whose type is unknown.
%
Our semantics \emph{conservatively} instantiates holes with concrete
values; thereby dynamically inferring the type of the input type
until the point the program goes wrong.
%
We prove that our procedure synthesizes \emph{general witnesses},
which means, intuitively, that if a witness is found for a given
ill-typed function, then \emph{for all} input types, there exist
inhabitants that can make the function go wrong.

Given a witness to a type error, the novice may still be at a loss.
%
The standard \ocaml\ interpreter and debugging infrastructure expect
well-typed programs, so they cannot be used to investigate \emph{how}
the witness causes the program to crash.
%
More importantly, the execution itself may be quite long and may contain
details not relevant to the actual error.

\paragraph{2. Visualizing Witnesses}
Our second contribution is a novel interactive visualization of the
execution of \ocaml\ programs well-typed or not~(\S~\ref{sec:interactive}).
%
We extend the semantics to also build a \emph{reduction graph}
which records all of the small-step reductions and the context
in which they occur.
%
The graph lets us build a visualization that shows a sequence of
steps from the source witness to the stuck term and allows the user
to interactively expand the  computation to expose intermediate steps,
by selecting an expression and choosing a traversal strategy.
%
The strategies include many of the standard debugging moves, \eg\
stepping \emph{forward} or \emph{into} or \emph{over} calls as well
stepping or jumping \emph{backward} to understand how a particular
value was created, while preserving a context of the intermediate
steps that allow the user to keep track of a term's provenance.

% This process of iterative expansion of the trace ensures that
% the context of intermediate steps is not lost; whereas traditional
% debuggers force one to move between statements, in our visualization
% you never lose sight of where you came from.

Obviously, the problem of finding witnesses is undecidable.
In fact, due to the necessarily conservative nature of static
typing, there may not even exist. Thus, our heuristic approach
is only useful if it can find \emph{compact} witnesses for
\emph{real-world} programs.

\paragraph{3. Evaluating Witnesses}
%
Our third contribution is an extensive evaluation of our approach
on two different sets of ill-typed programs obtained by instrumenting
compilers used in beginner's classes~(\S~\ref{sec:eval}).
%
The first is the \uwbench\ data set~\cite{lerner_searching_2007}
comprising \uwsize\ ill-typed programs.
%
The second is a new \ucsdbench\ data set, comprising \ucsdsize\
ill-typed programs.
%
We show that for both data sets, our technique is able to generate
witnesses for nearly 90\% of the programs, in under a second the
vast majority of cases.
%
Furthermore, we show that a simple interactive strategy yields
compact counterexample traces with fewer than 5 steps for 60\%
of the programs, and under 10 steps for 90\% of the programs.

In short, our results show that at least for beginners, ill-typed
programs \emph{do} go wrong, in the vast majority of cases, which,
in turn, opens the door to a completely different, dynamic way
to explain, understand, and appreciate the benefits of static typing.

%% Then, we formalize our algorithm and prove it sound
%% (\S~\ref{sec:searching-witness}).
%% %
%% Next, we extend the algorithm to collect a reduction graph and formalize
%% a set of traversals over the graph (\S~\ref{sec:explaining}).
%% %
%% Finally, we evaluate our search algorithm and trace generation on a
%% large set of real student programs (\S~\ref{sec:evaluation}).
%%
%% \paragraph{Outline}
%% We begin the rest of the paper with an overview of our search algorithm
%% and trace visualization (\S~\ref{sec:overview}).

% Then, we formalize our algorithm and prove it sound
% (\S~\ref{sec:searching-witness}), and evaluate it on a large set of real
% student programs (\S~\ref{sec:eval}).
% %
% Next, we extend the search algorithm to collect the reduction graph and
% formalize a set of traversals (\S~\ref{sec:explaining}), and measure the
% complexity of these traces on our benchmarks
% (\S~\ref{sec:evaluation-nanomaly}).
%
% Finally, we present the results of a preliminary user study on the full
% system (\S~\ref{sec:user-study}).

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: