### 2016-01-13 paper/pldi16-submission/response.txt
add first draft of response

We thank the reviewers for their enthusiastic comments, and will briefly
respond to the main critiques.

Formalism
---------

While we have implemented (and evaluated) our technique on the pure
subset of OCaml -- including algebraic datatypes, polymorphism, and type
annotations -- we restricted the formalism to a simpler language because
the additional constructs do not any major complexity to the technique.
We simply treat constructor application, case analysis, type annotation,
etc. as primitive operations like `+` and `if`; they narrow the partial
types of values and force instantiation. However, in light of the
reviews it is clear that this is not obvious to the reader, so we will
include an extended formal semantics in the next iteration of the paper.

We also thank Reviewer B in particular for pointing out the two bugs in
our proof. We have refactored the substitutions in the formalism to be
properly threaded through `narrow` and `E-Plus-Good`, and have changed
our definition of type compatibility to the usual one. While partial
types do grow monotonically, in Lemma 4 specifically we are concerned
with *any* value (and type) so the example with incompatible pair types
applies.

Representivity of the datasets
------------------------------

Between our dataset and the UW dataset, we have numerous instances of
writing higher-order functions and instantiating their arguments, as
well as two different expression evaluators. So we believe the
benchmarks are representative of the range of programs an undergraduate
student may have to write.

Pedagogic motivation for witnesses and traces
---------------------------------------------

We are also convinced of the importance of witnesses and traces for
debugging and did not intend this paper to specifically present evidence
in their favor, rather to show that generating witnesses is feasible and
*practical* for type errors. (This is also why we did not discuss
producing witnesses for other runtime errors like divide-by-zero.) We
will update the motivation to be clearer about the scope of the paper
and results.