### 2016-05-04 paper/icfp16-submission/rebuttal.txt
add list of extra papers to cite and discuss

We thank the reviewers for their thoughtful remarks, and will respond
to a few categories of concerns.

1. Related Work

We thank reviewers B and C for the pointers to extra related work. We
will greatly expand the section in the next revision of the paper to
include many of the following references.

- Haack and Wells, ESOP 2003
- Rahli et al., Tech Report 2010
- Schilling, TFP 2011
- Gast, IFL 2004
- Heeren et al., ICFP 2003
- Hage and Heeren, IFL 2006
- Christiansen, TFP 2014
- Serrano and Hage, ESOP 2016
- Chitil, ICFP 2001
- Stuckey et al., Haskell 2003
- Bayne et al., ICSE 2011
- Zhang et al., PLDI 2015

2. Bias in Type Variable Instantiation

In the `fac` example, asserting that `n` must be an int without looking
at the rest of the code is indeed biased, but any inference scheme will
be biased in one direction or another. Our bias is towards the
evaluation semantics of the language, other techniques are biased by
their traversal strategy (eg top-down vs bottom-up) or by various
heuristics (whether provided by the compiler or through generic
reasoning).

We would argue, however, (as in Section 5.4) that our bias is more
natural as it is based on a concrete instance of the program crashing.

3. Overloading / Type Classes

Overloaded operators can be problematic for our approach. In reviewer
C's variant of `fac`, our approach would get stuck at the `if n <= m`
test, because `n` and `m` are both unconstrained at this point (apart
from having the same type), so we do not know how to instantiate
them. This does not signify a type error of course, but our approach
would fail to discover a witness. We suspect this is relatively uncommon
in novice OCaml programs (there were no instances of this scenario in
either of our datasets), but the issue would surely be exacerbated if we
ported the approach to Haskell due to the prevalence of type classes.

We will add a paragraph describing this issue to the Discussion section,
and defer a treatment of overloading to future work.

4. Scaling to Real-World Code

Our benchmarks are small programs as they were drawn from homework
submissions. The UW dataset had programs generally in the 100-200 LoC
range, while the programs from our dataset were generally <50 LoC (we
had already sliced out a minimal program for each error in our dataset).
It is quite possible that our heuristic for compressing the trace based
on function calls would not suffice for larger programs, though there is
a broad literature on program slicing to draw from there.

5. Language Features

By "pure subset" of OCaml we mean polymorphic datatypes and records, and
we will clarify this in the text. We do not support references, modules,
or the object system.

It is true that we implicitly assume all types are inhabited in the
formalism, and that the generality theorem is weaker in the presence of
uninhabited types. We will clarify this in the text.