### 2016-05-04 paper/icfp16-submission/rebuttal.txt
revise related work rebuttal

We thank the reviewers for their thoughtful remarks, and will respond
to a few categories of concerns.

1. Related Work

We thank reviewers B and C for the pointers to extra related work.

Type error slicing shares a common goal with us of providing
more diagnostic information to the programmer, though the techniques
differ significantly. Error slicing is based on a static analysis of
the typing data, usually as a constraint system, and presents
multiple candidate error locations, possibly ranked according to
some heuristic. In contrast, our approach executes the program and
presents a concrete execution trace demonstrating the error. Gast's
IFL 2004 paper is perhaps the most similar to ours as he produces
a data flow that would violate the typing rules.

Customizable type error messages along the lines of Heeren 2003,
Christiansen 2015, and Serrano Mena 2016 also addresses the issue
of poor type errors. The difference with our work is more pronounced,
as these techniques often rely on a library author (or instructor)
with domain-specific knowledge to craft high-quality error messages,
whereas our approach is fully automatic but ignorant of the application
domain.

We might also add the algorithmic type debugging work of Chitil 2001
and Stuckey 2003, which presents the user with a series of queries
about the types of intermediate expressions to help isolate and
explain the error.

We will extend the Related Work section to reference all of these
techniques and papers.

2. Bias in Type Variable Instantiation

In the `fac` example, asserting that `n` must be an int without looking
at the rest of the code is indeed biased, but any inference scheme will
be biased in one direction or another. Our bias is towards the
evaluation semantics of the language, other techniques are biased by
their traversal strategy (eg top-down vs bottom-up) or by various
heuristics (whether provided by the compiler or through generic
reasoning).

We would argue, however, (as in Section 5.4) that our bias is more
natural as it is based on a concrete instance of the program crashing.

3. Overloading / Type Classes

Overloaded operators can be problematic for our approach. In reviewer
C's variant of `fac`, our approach would get stuck at the `if n <= m`
test, because `n` and `m` are both unconstrained at this point (apart
from having the same type), so we do not know how to instantiate
them. This does not signify a type error of course, but our approach
would fail to discover a witness. We suspect this is relatively uncommon
in novice OCaml programs (there were no instances of this scenario in
either of our datasets), but the issue would surely be exacerbated if we
ported the approach to Haskell due to the prevalence of type classes.

We will add a paragraph describing this issue to the Discussion section,
and defer a treatment of overloading to future work.

4. Scaling to Real-World Code

Our benchmarks are small programs as they were drawn from homework
submissions. The UW dataset had programs generally in the 100-200 LoC
range, while the programs from our dataset were generally <50 LoC (we
had already sliced out a minimal program for each error in our dataset).
It is quite possible that our heuristic for compressing the trace based
on function calls would not suffice for larger programs, though there is
a broad literature on program slicing to draw from there.

5. Language Features

By "pure subset" of OCaml we mean polymorphic datatypes and records, and
we will clarify this in the text. We do not support references, modules,
or the object system.

It is true that we implicitly assume all types are inhabited in the
formalism, and that the generality theorem is weaker in the presence of
uninhabited types. We will clarify this in the text.