### 2017-03-05 paper/jfp-submission/conclusion.tex
let's see what the JFP format looks like..

\section{Conclusion}
\label{sec:conclusion}
In this paper we have demonstrated how to find \emph{dynamic} witnesses
to \emph{static} type errors.
%
Dynamic witnesses provide the user with more information than static
error messages, and enable the use of standard debugging techniques
to explore \emph{how} the program ``goes wrong''.
%
Our witness search produces general witnesses, precluding the
possiblity of a valid typing for the program.
%
We have shown a novel interactive visualization of the execution of
functional programs, which can be used to \emph{show} a novice how a
witness makes their program crash.
%
Our evaluation shows that our search is effective at finding witnesses
with a 87\% success rate, and that our jump-compressed traces
effectively hide boring details of the execution, with 90\% of the
benchmarks containing less than 10 jumps.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: