### 2017-04-24 paper/jfp-submission/safe.tex
discuss speculative insantiation

\subsubsection{How Safe are the ``Safe'' Programs?}
\label{sec:how-safe}

An immediate question arises regarding that 13--16\% of programs for
which we could not synthesize a witness:
%
are they \emph{actually} safe (\ie is the type system being too conservative),
%
or did \toolname simply fail to find the witness?

\ES{Need better organization for the groupings of programs, maybe a pie chart to guide us?}

To answer this question, we investigated the 720 \ucsdbench programs for
which we failed to find a witness.
%
We found that for 247 programs (roughly a third of the failures, or 5\%
of all programs), \toolname got stuck when it tried to compare two holes.
%
\ocaml provides polymorphic equality and comparison operators,
overloading them for each type.
%
While extremely convenient to use, they pose a challenge for \toolname's
combination of execution and inference.
%
Consider, for example
%
\begin{code}
  let bad x y =
    if x < y then
      1 + true
    else
      0
\end{code}
%
When given @bad@ as input, \toolname will generate two fresh holes
$\nu_1[\alpha_1]$ and $\nu_2[\alpha_2]$ and proceed directly into the
@x < y@ comparison.
%
We cannot (yet) instantiate either hole because we have no constraints
on the $\alpha$s (we know they must be equal, but nothing else), and
furthermore we do not know what constraints we may encounter later on in
the program.
%
Thus, we cannot perform the comparison and proceed, and must give up our
search for a witness, even though one obviously exists, any pair of |x|
and |y| such that |x < y| is true.

Extending \toolname with support for symbolic execution would alleviate
this issue, as we could then begin symbolically executing the program
until we learn how to instantiate |x| and |y|.
%
Alternatively, we could \emph{speculatively} instantiate both |x| and
|y| with some arbitrary type, and proceed with execution until we
discover a type error.
%
This speculative instantiation is of course unsound, we would have to
take care to avoid reporting frivolous type errors that were caused by
such instantiations.
%
We would need to track which holes were instantiated speculatively in
order to distinguish type errors that would have happened regardless, as
in the above @bad@ function, from type errors that were the fault of our
instantiation.

Further, suppose that our speculative instantiation induces a frivolous
type error.
%
For example, suppose we are given
%
\begin{code}
  let bad x y =
    if x < y then
      x *. y
    else
      0.
\end{code}
%
and choose to (speculatively) instantiate @x@ and @y@ as @int@s and proceed
down the ``true'' branch.
%
We will quickly discover this was the wrong choice, as they are immediately
narrowed to @float@s.
%
We must now backtrack and try a different instantiation, but we no
longer need to choose one at random.
%
Since our instantiation was speculative, we can still view @x@ and @y@
\emph{morally} as holes, and thus treat the @*.@ operator as a normal
narrowing point with two holes.
%
This tells us that the \emph{correct} instantiation was in fact @float@,
and we can then proceed as normal from the backtracking point with a
concrete choice of @float@s.

Thus, it appears that speculative instantiation of holes may be a
useful, lightweight alternative to symbolic execution for our purposes.

The remaining 473 programs for which we could not produce a witness did
not admit such an automatic diagnosis, so we selected a random sample of
50 programs and manually searched for a witness.
%
We categorized the programs into four groups.

\paragraph{Witness Exists}
%
We found only eight cases (or 16\% of the samples) where a witness existed,
but \toolname was unable to discover it.
%
Five involved synthesizing a \emph{pair} of specially-crafted inputs
that would result in the function returning values of incompatible
types.
%
The other three required synthesizing an input that would trigger a
particular path through the program, and would likely have been caught
by symbolic execution.
%
% Overall, it appears that our random instantiation of holes is well-suited
% to finding type errors in student programs.
%
%We categorized the remaining 42 failures into three groups.

\paragraph{Dead Code}
%
Six programs (or 12\% the samples) contained type errors that were
unreachable by any inputs, often due to overlapping patterns in a
@match@ expression.
%
While technically safe, dead code is generally considered a ``code
smell''; students would likely benefit from a warning in this case.

\paragraph{Bad Function Calls}
Fourteen programs (or 28\% of the samples) contained a function call
where the student had supplied incorrect inputs, yet no error was
triggered.
%
Consider, for example the following |assoc| function, which
looks up a key in an \emph{association list} and returns a default if
the key cannot be found.
%
\begin{code}
  let rec assoc (d, k, l) = match l with
    | (ki, vi)::tl ->
       if ki = k then
         vi
       else
         assoc (d, k, tl)
    | _ -> d

  let _ = assoc ([], 123, [(123, "sad"); (321, "happy")])
\end{code}
%
The student's definition of |assoc| is correct, but \ocaml rejects their
subsequent call because the default value |[]| is incompatible with the
|string| values in the list.
%
In this particular call, the key |123| is in the list, so the default
will not be used (and even if it were, no error would be triggered) and
\ocaml's complaint is moot.
%
Of course, \ocaml cannot be expected to know that this particular call
is safe, its type system is not sophisticated enough to express the
necessary conditions.

We note that all of the programs in this category appear to come from
interactions with the \ocaml top-level, \ie the student defined a
function and is now experimenting with it.
%
These calls are thus particularly harmless, indeed they may even be
\emph{beneficial}, as the student may next try looking up a key that
\emph{does not} exist, and realize that the returned values are
incompatible.

\paragraph{Non-Parametric Function Type}
%
Twenty-two programs (or 44\% of the samples) lack a witness in our
semantics due to our non-parametric $\tfun$ type for functions.
%
Recall that our goal is to expose the runtime errors that would have
been prevented by the type systems.
%
At runtime, it is always safe to call a function, thus we give functions
a simple type $\tfun$ that says they may be applied, but says nothing
about their inputs or outputs.
%
But consider the following @clone@ function, which is supposed to
produce a list containing @n@ copies of the input @x@.
%
% Seven programs violated the \emph{occurs check} with cyclic typing
% constraints like @'a = 'a list@, for example the following @clone@
% function.
%
\begin{code}
  let rec clone x n =
    if n > 0 then
      clone [x] (n - 1)
    else
      []
\end{code}
%
Unfortunately, the student instead constructs an @n@-level nested list
containing a single @x@.
%
The \ocaml compiler rejects this program because the recursive call to
@clone@ induces a cyclic typing constraint @'a = 'a list@, capturing the
fact that each call increases the nesting of the list.
%
\toolname fails to catch this because we do not track the types of the
inputs to @clone@.

We note, however, that @clone@ cannot go wrong; it is perfectly safe to
repeatedly enclose a list inside another (disregarding the fact that the
nested list is never returned).
%
Still, such a function would be very difficult to \emph{call} safely, as
the programmer would have to reason about the dependency between the
input @n@ and the nesting of the output list, which cannot be expressed
in \ocaml's type system.
%
Thus, is it not particularly satisfying that \toolname fails to produce
a witness here; a possible solution could be to track the types of the
inputs, and demonstrate to the user how they change between recursive
calls.

To summarize, our investigation suggests that the majority of programs
(88\%) for which we fail to find a witness do not, in fact, admit a
witness.
%
Overall, it appears that our random instantiation of holes is well-suited
to finding type errors in student programs.
%
The remaining programs were all cases where \ocaml's type system was
overly conservative.
%
Of course, the conservatism is somewhat justified as each case pointed
to a ``code smell'' or difficult-to-use API;\ it would be interesting to
investigate how demonstrate these issues in a similar manner.

% \begin{enumerate}
% \item 7 infinite type. program doesn't go wrong, but could search for
%   context that would go wrong
% \item 15 monomorphic functions. recall we abstract a function's type as
%   $\tfun$ rather than $\alpha \to \beta$. (may also admit witness by
%   searching for context?)
% \item 6 dead code. type error occurs in unreachable code, ``deficiency''
%   of the type system.
% \item 14 bad call. function is well-typed, student calls makes ill-typed
%   call, but no error triggered. nothing to be done, as we do not control
%   call-site.
% \end{enumerate}