### 2017-12-10 paper/jfp-submission/response.txt
start responding to reviews

We thank the reviewers for their detailed comments and suggestions.

## Questions about the user study

Reviewer 1
> I was not completely convinced by the metrics used for the statistical
> analysis; I would have found it more convincing to measure the speed
> at which an incorrect program may be fixed by a student sitting in front
> of a machine. That said, some statistical analysis is better than none,
> and I am greatful that the authors attempted some evaluation.

Reviewer 2
> 7. why did you do a user study as part of an exam, and not as an
> independent study? I don't think this is a good idea.

We agree that it would be preferable to have timing data in addition to
correctness data. We did attempt a study that would have collected
timing data, but unfortunately were not able to recruit enough
participants to draw any conclusions. We have updated the text in the
"Constuct" subsection of the "Threats to Validity" to mention lack of
timing data as a threat.

Reviewer 2
> 2. you compare only to OCAML in your student user study, which seems
> to be to be an easy target. Why not Sherrloc or Mycroft which at least
> get the location right more often? If OCAML much more often gives the
> wrong location, the message is not likely to be helpful. Indeed, I
> would expect you to only compare the systems for those cases that both
> NANOMALY and OCAML blame the same location.

Reviewer 2
> 3. does your tool help students to learn about the type system, or is
> it simply a better crutch to help them along. Did you investigate this
> at all?

These are valid concerns, we have updated the "Threats to Validity" to
mention them.

Reviewer 3
> The discussion of inter-rater reliability is strange. Why do we need
> to know if the exams are graded similarly, except to check that UCSD
> students are fairly graded? Also, why is IRR so low, given that
> these are questions with a correct answer? Given that there are only
> 60 exams in question, having a single evaluator with a clear rubric
> would eliminate both this potential source of inconsistency as well
> as the need to discuss IRR at all.

Our use of IRR is intended to account for potential bias in the
raters. One such form of bias (conscious or not) could be bias in favor
of NanoMaLy since it is new and interesting, and the raters (even if
they are not directly connected to the project) may want to see it
succeed. Due to the graphical nature of NanoMaLy's traces it is not easy
to hide whether a given answer was assisted by OCaml or NanoMaLy from
the raters. A further form of potential bias is inconsistency in
evaluating the student answers, which were long-form, English
explanations of the error/fix. There is a surprising amount of
subjectivity involved in judging, based on a couple sentences, whether
the student truly understood the error.

For these reasons we feel that our use of IRR is justified to lend
credibility to the results.

(We have not added this discussion to the paper as the IRR test is
really meant to be a preliminary "sanity-check", but would be happy to
if the reviewers feel it would be helpful.)

## Reviewer 1

> The tool is restricted to a small subset of OCaml, in particular it does
> not support side-effects. While many FP courses begin by focusing on
> pure programs only, there are some intringuing type errors that students
> face only when working with effectful programs, e.g. forgetting to apply
> a function to the unit argument. I speculate that it would be /not/ so
> easy to generalize the present work to deal with side effects, because
> of the need to maintain the state and report it to the user.

TODO

> Random-test generation would likely fail to produce counter-examples
> on slightly more complex programs such as red-black trees; for such
> programs, more advanced techniques such as constraint-based test
> generation are probably required. I was not able to quite understand
> whether the approach presented in the paper could be extended to
> accomodate constraint-based test generation.

TODO

> page 23, regarding section "witness exists", I got puzzled. The study
> there is on 50 programs. It was found that 2% of the programs admit a
> witness. I deduce this must correspond to one single program. But
> then, the next sentence tells that "slightly over half of these
> programs involved had some particular property, while the rest had
> some other property."  There semms to be some inconcistency there.

Indeed, the wording here was confusing. The percentages in this section
are presented as a fraction of *all* programs, not just the random
sample of failing programs. We have updated the text at the start of 5.3
to explain this.

## Reviewer 2

> 1. you do not address the issue of bias, the fact that locations
> towards the end of the program are more likely to be crashed at. This
> is similar to the left-to-right-bias of algorithm W, although in your
> case the left-to-rightness is in the trace, not the syntax tree (as
> addressed for that case in work by Hage and Heeren (VODCA 2008) and
> the work of Erwig and Chen).  Does your implementation allow general
> datatypes so that I can have case with many arms?

TODO

> 5. your approximation of functions to fun is rather strong, and it is
> unclear to me how well that works. I guess since all executions start
> from the beginning of the programs keeps everything rather concrete,
> so that you do not need to invent functions. But I am not sure, and I
> would like to see a thorough discussion. I am fine with novice
> programs being short and simple since that is quite likely to be the
> case when they resort to Nanomaly, but when you do that, it should be
> clear that you do.  What may be related to this is the idea to deal
> with Currying. What about functions like foldr of which the result is
> a type variable? I know that some static work on error diagnosis
> sometimes runs into a problem of some sort and since you work at the
> dynamic level this may not be the case for you. But do say and explain
> so.

TODO

> 1. explain how things would be if applied to other functional
> languages, in particular Haskell. What about laziness? What about
> overloaded numerals?  It seems to me that these are quite
> problematic. For example, the 0 in your fac program will be Num a =>
> a.

TODO

> 2. what is the limit to the size of programs you can reasonably
> handle?  What if we have a large program, is it possible to have a
> different starting point? So does a trace have to start at the start
> of the program? I guess this is important since if you start with a
> higher order function then you need to dynamically invent a function,
> and then your approximation might well be in the way. This is
> something I am really curious about.  Now you have 5% Non-parametric
> function types. Does that change in this situation?

TOOD